---
title: "Lab 7: Murder in SQL City"
author: "Michael Torrisi"
date: today
format: html
execute: 
  echo: true
  error: true
  embed-resources: true
  output: false
code-tools: true
toc: true
theme: default
fontcolor: black
linkcolor: blue
monobackgroundcolor: rgb(199, 221, 235)
---

For this lab, you will be joining and filtering related data sets to solve a murder mystery!

# Part 1: GitHub Workflow

At the top of the document (in the YAML) there is an `author` line that says 
`"Your name here!"`. Change this to be your name and save your file either by
clicking on the blue floppy disk or with a shortcut (command / control + s). 

Be sure to [commit the files to your repo](https://happygitwithr.com/existing-github-first#stage-and-commit). 

Let's get started!


# Part 2: Some Words of Advice

-   Set chunk options carefully.

-   Make sure you don't print out more output than you need.

-   Make sure you don't assign more objects than necessary---avoid "object junk"
in your environment.

-   Make your code readable and nicely formatted.

-   Think through your desired result **before** writing any code.

# Part 3: Finding the Killer

Northwestern University's Knight Lab wanted to help sharpen users' database
skills, so they created a murder mystery. Can you solve this crime in SQL City??

The relational data you will be working with contains tables with different
pieces of information pertinent to the crime - people, social media check-ins,
driver's licenses, crime scene reports, police interviews, and more!


## Access the Data

This code chunk will read in **all** of the tables of data for you. Don't modify or
remove this! Take some time to look at each file type so that 

```{r}
#| message: true
#| warning: false
#| label: read-in-data

library(tidyverse)

# If purrr is not detected, install the package
if (!"purrr" %in% installed.packages()) install.packages("purrr")

source("https://raw.githubusercontent.com/jcanner/stat_210_2025_website/main/labs/instructions/lab-7-setup.R")
```

## Solve the Crime

### Crime Scene Report

Detective Wickham reaches out to you...

> A crime has taken place and I need your help! There was a murder in SQL City sometime on January 15, 2018. Could you retrieve the crime scene report from the police department’s database and follow the clues to find the person responsible for the murder?!


**Step 1: Find the police report in `crime_scene_report`. Then used the information in the report to move on to the next data source based on the information learned.**

#### Finding Relevant Report
```{r}
#| label: Finding Relevant Report
# Filtered police reports to find one with relevant characteristics.
crime_scene_report %>% 
  filter(date == 20180115 &
           type == "murder" &
           city == "SQL City") %>% 
  select(description)
```

> Security footage shows that there were 2 witnesses. The first witness lives at the last house on "Northwestern Dr". The second witness, named Annabel, lives somewhere on "Franklin Ave".

**Next Steps: Follow the evidence to the person responsible for the murder, building a report as you go.**
There are accomplices, some knowingly and some unknowingly, but there is only one mastermind.


Solve the murder mystery, showing **all of your work in this document**. Your
document and code must be well organized, easy to follow, and reproducible.

+ Use headers and written descriptions to indicate what you are doing.
+ Use `dplyr` verbs and `join` functions rather than just looking through the
tables manually. Functions from `stringr` and `lubridate` will also be useful.
+ Use good code formatting practices.
+ Comment your code.
+ Cite any external sources you use to solve the mystery.

#### Renaming Variables
```{r}
#| label: Renaming Variables
# Renaming variables to facilitate data joins.
interview <- interview %>% 
  rename(id = person_id)

get_fit_now_member <- get_fit_now_member %>% 
  rename(gym_id = id,
         id = person_id)

facebook_event_checkin <- facebook_event_checkin %>% 
  rename(id = person_id)

drivers_license <- drivers_license %>% 
  rename(license_id = id)
```

#### Joining Data
```{r}
#| label: Joining Data
# Joining all the dataframes into one central dataframe.
full_info <- person %>% 
  left_join(income,
            by = "ssn") %>% 
  left_join(interview,
            by = "id") %>% 
  left_join(get_fit_now_member,
            by = "id") %>% 
  left_join(facebook_event_checkin,
            by = "id") %>% 
  left_join(drivers_license,
            by = "license_id") %>% 
  select(-name.y) %>% 
  rename(name = name.x) %>% 
  distinct(id,
           .keep_all = TRUE)
```



#### First Witness & Statement
```{r}
#| label: First Witness
# Finding the name and statement of the first witness.
full_info %>% 
  filter(address_street_name == "Northwestern Dr") %>% 
  arrange(desc(address_number)) %>% 
  slice_head(n = 1) %>% 
  select(name,
         transcript)
```

> Morty's Statement: I heard a gunshot and then saw a man run out. He had a "Get Fit Now Gym" bag. The membership number on the bag started with "*48Z*". Only gold members have those bags. The man got into a car with a plate that included "*H42W*".

#### Second Witness & Statement
```{r}
#| label: Second Witness
# Finding the name and statement of the second witness.
full_info %>% 
  filter(address_street_name == "Franklin Ave") %>% 
  mutate(first_name = substring(name, 1, 7)) %>% 
  filter(first_name == "Annabel") %>% 
  select(name,
         transcript)
```

> Annabel's Statement: I saw the murder happen, and I recognized the killer from my gym when I was working out last week on *January the 9th*.

#### Gym Membership Numbers 
```{r}
#| label: Gym Membership Numbers
# Finding gym membership id numbers that fit the witness's reports.
get_fit_now_check_in %>% 
  mutate(first3 = substring(membership_id, 1, 3)) %>% 
  filter(first3 == "48Z" &
         check_in_date == 20180109) %>% 
  select(membership_id)
```

#### Relevant Gym Members' Names & Statements
```{r}
#| label: Relevant Gym Members' Names & Statements
# Finding the names and statements of gym members whose id numbers fit the witness's reports.
full_info %>% 
  filter(!is.na(gym_id)) %>% 
  filter(id %in% c(61749, 62783) |
         gym_id %in% c("48Z7A", "48Z55")) %>% 
  select(name,
         transcript)
```

> Jeremy Bowers statement: "I was hired by a woman with *a lot of money*. I don't know her name but I know she's around 5'5" (*65*") or 5'7" (*67*"). She has red hair and she drives a Tesla *Model S*. I know that she attended the *SQL Symphony Concert* 3 times in December 2017."

#### Final Suspect
```{r}
#| label: Final Suspect
#| output: true
# Finding our final suspect, based on information obtained from Jeremy Bowers' statement.
full_info %>% 
  filter(gender == "female" &
           height >= 65 &
           height <= 67 &
           hair_color == "red" &
           car_model == "Model S" &
           event_name == "SQL Symphony Concert") %>% 
  select(name)
```


::: callout-note
## And the final suspect is...

*Miranda Priestly*
:::
